\LaTeX\ Author Guidelines for ICCV Proceedings

### 2021-03-16 chapters/2_rel_work.tex
Shorten text

\textbf{Proposal-free instance segmentation methods} adopt a bottom-up approach by directly grouping pixels into instances. Recently, there has been a growing interest in such  methods that do not involve object detection because, in certain types of data, object instances cannot be approximated by bounding boxes \cite{kirillov2017instancecut,bai2017deep}. 
% For example, the approach proposed in \cite{kirillov2017instancecut} uses a combinatorial framework for instance segmentation, 
% SGN \cite{liu2017sgn} sequentially group pixels into lines and then instances;
% whereas a watershed transform is learned in \cite{bai2017deep} by also predicting its gradient direction. 
% whereas the template matching \cite{uhrig2016pixel} deploys scene depth information.
Some use metric learning to predict high-dimensional associative pixel embeddings that map pixels of the same instance close to each other \cite{lee2019learning,fathi2017semantic,newell2017associative,de2017semantic}
% , while mapping pixels belonging to different instances further apart \cite{lee2019learning,fathi2017semantic,newell2017associative,de2017semantic}. % kulikov2018instance
and then retrieve final instances by applying a clustering algorithm \cite{kong2018recurrentPix}.
% , like in the end-to-end trainable mean-shift pipeline of \cite{kong2018recurrentPix}. 
Other recent methods let the model predict the relative coordinates of the instance center \cite{neven2019instance,cheng2019panopticdeeplab} or, given a pixel $(x,y)$, they train a model to generate the mask of the instance located at $(x,y)$ \cite{sofiiuk2019adaptis}.

>>>>

\textbf{Proposal-free instance segmentation methods} adopt a bottom-up approach by directly grouping pixels into instances. In the last years, there has been a growing interest in such  methods that do not involve object detection because, in certain types of data, object instances cannot be approximated by bounding boxes \cite{kirillov2017instancecut,bai2017deep}. 
% For example, the approach proposed in \cite{kirillov2017instancecut} uses a combinatorial framework for instance segmentation, 
% SGN \cite{liu2017sgn} sequentially group pixels into lines and then instances;
% whereas a watershed transform is learned in \cite{bai2017deep} by also predicting its gradient direction. 
% whereas the template matching \cite{uhrig2016pixel} deploys scene depth information.
Some use metric learning to predict high-dimensional associative pixel embeddings that map pixels of the same instance close to each other \cite{lee2019learning,fathi2017semantic,newell2017associative,de2017semantic}
% , while mapping pixels belonging to different instances further apart \cite{lee2019learning,fathi2017semantic,newell2017associative,de2017semantic}. % kulikov2018instance
and then retrieve final instances by applying a clustering algorithm \cite{kong2018recurrentPix}.
% , like in the end-to-end trainable mean-shift pipeline of \cite{kong2018recurrentPix}. 
Other recent methods let the model predict the relative coordinates of the instance center \cite{neven2019instance,cheng2019panopticdeeplab} or, given a pixel $(x,y)$, they train a model to generate the mask of the instance located at $(x,y)$ \cite{sofiiuk2019adaptis}.

### 2021-03-16 chapters/3_algorithm.tex
Shorten text

\textbf{Linkage criteria and hierarchical trees} -- 
% We also denote as $S_u$ the cluster associated with node $u$ \TODO{Needed?}.
% We call two clusters $S,S'$ \emph{adjacent} if there exists at least one edge ${e_{ts}\in E}$ connecting a node $t\in S_u$ to a node $s\in S_v$. 
Let the interaction $\interact(S,S')\in\mathbb{R}$ between two clusters $S,S'$ be defined as a function, named \emph{linkage criterion}, depending on the weights of \emph{all} edges connecting clusters $S$ and $S'$, i.e. ${(S \times S') \cap E}$. 
The linkage criteria tested in this article are listed and defined in Table \ref{tab:linkage-criteria}.
A \emph{dendrogram} $T$ is a rooted binary tree\footnote{In general, one could look at trees that are not binary. However, the algorithms discussed in this paper always generate binary hierarchical trees, so nothing would be gained by this generalization.} representing the merging order of an agglomerative algorithm, such that the leaves of the tree are in one-to-one correspondence with $V$. 
Let $T_{\mathrm{R}},T_{\mathrm{L}}\subset T$ denote the subtrees rooted at the two children of the root node in $T$.
For any two leaves $u,v \in V$, let $T[u \vee v]$ be the subtree rooted at the least common ancestor $(u \vee v)\in T$ of nodes $u$ and $v$ (furthest from the root), and let \texttt{leaves}$(T[u \vee v])\subseteq V$ be the set of leaves of this subtree. 
% \TODO{Here we talk about the algorithm already, so we could move it to next section and the flow could be better. But we need a tree in the alg. So we better do this later}
% Algorithm \ref{main_alg} returns a binary tree denoted by $T^*$.
Given an agglomerative algorithm with merging tree $T$, let $h_T:T \rightarrow \mathbb{N}$ denote the \emph{dendrogram-height} of each $(u\vee v)\in T$, which is defined as the iteration number at which nodes $u,v\in V$ were merged by the algorithm (see example in Fig.~\ref{fig:intro_figure}a). We also define $\treeHeight(u,v)$ as the signed interaction $\interact{}(S,S')$ between the two clusters $S,S'$ that were merged at iteration $h_T(u\vee v)$: 
% Then, given a binary tree $T$ and a linkage criterion $\interact$, we can associate a \emph{signed similarity} function $\treeHeight$ to each pair of distinct original nodes $u\neq v$, with $u, v \in V$:
\begin{equation}\label{eq:def_dendr_interact}
\treeHeight(u,v) \equiv \interact{} \big( \text{\texttt{leaves}}(T_{\mathrm{R}}[u \vee v]), \text{\texttt{leaves}}(T_{\mathrm{L}}[u \vee v]) \big)
\end{equation}
% \OPTIONAL{In other words, $h_T(u,v)$ and $\treeHeight(u,v)$ represent the \emph{dendrogram-height} and \emph{signed-interaction} associated to each $u \vee v$ node in the tree $T$ (see toy example in \TODO{Fig.~1}).}

@@

In phase one, \algname{} selects the pair of clusters with the highest absolute interaction $|\interact(S, S')|$, so that the most attractive and the most repulsive pairs are analyzed first. If the interaction is repulsive and the algorithm option \emph{addCannotLinkConstraints} is \texttt{True}, then the two clusters are constrained so that their members can never merge in subsequent steps of phase one. If the interaction is attractive, then the clusters are merged, provided that they were not previously constrained. 
After each merge, the interaction between the merged cluster and its neighbors is updated according to one of the linkage criteria $\interact(S, S')$ listed in Table \ref{tab:linkage-criteria}. Phase one terminates when all the remaining clusters are either constrained or share repulsive interactions. Note that, on unsigned graphs, in phase one \algname{} merges all nodes in a single cluster and is then equivalent to a standard hierarchical clustering algorithm.

@@

\paragraph{AbsMax Linkage} -- This linkage method is also specific to signed graphs, since on unsigned graphs it would be equivalent to single linkage. Here, we prove that the Mutex Watershed Algorithm \cite{wolf2018mutex} can be seen as an agglomerative algorithm with AbsMax linkage (proofs of the following three propositions are given in Appendix \ref{sec:proposition_proofs}):
\begin{restatable}{prop}{absmaxmutex}
\label{prop:absmax_mutex}
The \algname{} Algorithm \ref{main_alg} with AbsMax linkage (see def. in Table \ref{tab:linkage-criteria}) returns the same final clustering $\Pi^*_{\mathrm{AbsMax}}$ whether or not cannot-link constraints are introduced. Also, $\Pi^*_{\mathrm{AbsMax}}$ is equal to the clustering returned by the Mutex Watershed Algorithm (MWS) \cite{wolf2018mutex}, which has empirical complexity $\mathcal{O}(N \log N)$. 
\end{restatable}
% \begin{prop} 
% \end{prop}

\paragraph{Average, Single, and Complete Linkage} -- These three linkage criteria have been throughly studied on unsigned graphs, but never on signed graphs. In recent independent related work \cite{chehreghani2020hierarchical}, the authors prove that applying these three linkage methods to a signed graph is equivalent to applying them to the unsigned graph obtained by shifting all edge weights by a constant.  Here, we prove which of the algorithms studied here is ``intrinsically'' signed and do not have this invariance-property:
\begin{restatable}{prop}{invariantAlgs}
\label{prop:weight_shift_invariant}
We call an agglomerative algorithm ``weight-shift invariant'' if the structure of the merging tree $T$ returned by the algorithm is invariant w.r.t. a shift of all edge weights $w_e$  by a constant $\alpha\in \mathbb{R}$. Among the versions of \algname{} studied in this paper, only hierarchical clustering using Average (HC-Avg), Single (HC-Single), and Complete linkage (HC-Complete) are weight-shift-invariant (see algorithms included in the green box in Table~\ref{tab:linkage-criteria}).
\end{restatable}
\noindent To our knowledge, Average and Single linkage methods have never been combined with cannot-link constraints on signed graphs, so we name these algorithms \emph{HCC-Avg} and \emph{HCC-Single}. On the other hand, we prove that enforcing constraints with a Complete linkage does not change the final clustering (Lemma \ref{lemma:absMax_and_complete_property}, in Appendix).
% , we prove that with Complete linkage the agglomeration order does not change whether constraints are introduced or not.

@@

\paragraph{Algorithms defining an ultrametric} -- The connection between agglomerative algorithms and ultrametrics\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.} is well known \cite{johnson1967hierarchical,milligan1979ultrametric}. Usually, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. Here, we generalize this concept to signed graphs. To do so, we first map the signed interaction $\treeHeight{}$ defined in Eq.~\ref{eq:def_dendr_interact} to positive ``pseudo-distances'' $d_{T}:V \times V \rightarrow \mathbb{R}^{+}$:
\begin{align}\label{eq:UM_def}
d_{T}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv  \epsilon + \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v')
\end{align}
and where $\epsilon > 0$. 
% By definition, we have that $d_{T}(u,v)=d_{T}(v,u)$ and $d_{T}(u,u)= 0$. 
We then prove the following proposition:
\begin{restatable}{prop}{secondUltraMetricProperty}
\label{prop:ultraMetric2}
% Consider the binary rooted tree $T^*$ returned by the \TODO{all variants in Table 1} \algname{} Algorithm~\ref{main_alg}, given a choice of linkage criteria $\interact{}$ and the use (or not) of CannotLinkConstraints. 
Among the algorithms included in the \algname{} framework (see Table \ref{tab:linkage-criteria}), only Mutex Watershed and hierarchical clustering with Average (HC-Avg), Single (HC-Single) and Complete linkage (HC-Complete) define an ultrametric $(V, d_{T^*})$, where $d_{T^*}$ is defined in Eq.~\ref{eq:UM_def} and $T^*$ is the tree returned by the \algname{} Algorithm~\ref{main_alg}.
\end{restatable}

Now that we have defined the algorithms and managed to characterize them from a theoretical point of view, we move to empirical evaluation.

>>>>

\textbf{Linkage criteria and hierarchical trees} -- 
% We also denote as $S_u$ the cluster associated with node $u$ \TODO{Needed?}.
% We call two clusters $S,S'$ \emph{adjacent} if there exists at least one edge ${e_{ts}\in E}$ connecting a node $t\in S_u$ to a node $s\in S_v$. 
Let the interaction $\interact(S,S')\in\mathbb{R}$ between two clusters $S,S'$ be defined as a function, named \emph{linkage criterion}, depending on the weights of \emph{all} edges connecting clusters $S$ and $S'$, i.e. ${(S \times S') \cap E}$ where $\times$ denotes the outer product. 
The linkage criteria tested in this article are listed and defined in Table \ref{tab:linkage-criteria}.
A \emph{dendrogram} $T$ is a rooted binary tree\footnote{In general, one could look at trees that are not binary. However, the algorithms discussed in this paper always generate binary hierarchical trees, so nothing would be gained by this generalization.} representing the merging order of an agglomerative algorithm, such that the leaves of the tree are in one-to-one correspondence with $V$ and each node of the tree represents a merge between two clusters. 
Let $T_{\mathrm{R}},T_{\mathrm{L}}\subset T$ denote the subtrees rooted at the two children of the root node in $T$.
For any two leaves $u,v \in V$, let $T[u \vee v]$ be the subtree rooted at the least common ancestor $(u \vee v)\in T$ of nodes $u$ and $v$ (furthest from the root), and let \texttt{leaves}$(T[u \vee v])\subseteq V$ be the set of leaves of this subtree. 
% \TODO{Here we talk about the algorithm already, so we could move it to next section and the flow could be better. But we need a tree in the alg. So we better do this later}
% Algorithm \ref{main_alg} returns a binary tree denoted by $T^*$.
Given an agglomerative algorithm with merging tree $T$, let $h_T:V \times V \rightarrow \mathbb{N}$ denote the \emph{dendrogram-height} of each $(u\vee v)\in T$, which is defined as the iteration number at which nodes $u,v\in V$ were merged by the algorithm (see example in Fig.~\ref{fig:intro_figure}a). We also define $\treeHeight(u,v)$ as the signed interaction $\interact{}(S,S')$ between the two clusters $S,S'$ that were merged at iteration $h_T(u, v)$: 
% Then, given a binary tree $T$ and a linkage criterion $\interact$, we can associate a \emph{signed similarity} function $\treeHeight$ to each pair of distinct original nodes $u\neq v$, with $u, v \in V$:
\begin{equation}\label{eq:def_dendr_interact}
\treeHeight(u,v) \equiv \interact{} \big( \text{\texttt{leaves}}(T_{\mathrm{R}}[u \vee v]), \text{\texttt{leaves}}(T_{\mathrm{L}}[u \vee v]) \big)
\end{equation}
% \OPTIONAL{In other words, $h_T(u,v)$ and $\treeHeight(u,v)$ represent the \emph{dendrogram-height} and \emph{signed-interaction} associated to each $u \vee v$ node in the tree $T$ (see toy example in \TODO{Fig.~1}).}

@@

In phase one, \algname{} selects the pair of clusters with the highest absolute interaction $|\interact(S, S')|$, so that the most attractive and the most repulsive pairs are analyzed first. If the interaction is repulsive and the algorithm option \emph{addCannotLinkConstraints} is \texttt{True}, then the two clusters are constrained so that their members can never merge in subsequent steps of phase one. If the interaction is attractive, then the clusters are merged, provided that they were not previously constrained. 
After each merge, the interaction between the merged cluster and its neighbors is updated according to one of the linkage criteria $\interact(S, S')$ listed in Table \ref{tab:linkage-criteria}. Phase one terminates when all the remaining clusters are either constrained or share repulsive interactions. Note that, on unsigned graphs, in phase one all nodes are merged into a single cluster and \algname{} is then equivalent to a standard hierarchical clustering algorithm.

@@

\paragraph{AbsMax Linkage} -- This linkage method is also specific to signed graphs, since on unsigned graphs it would be equivalent to single linkage. Here, we prove that the Mutex Watershed Algorithm \cite{wolf2018mutex} can be seen as an agglomerative algorithm with AbsMax linkage (proofs of the following three propositions are given in Appendix \ref{sec:proposition_proofs}):
\begin{restatable}{prop}{absmaxmutex}
\label{prop:absmax_mutex}
The \algname{} Algorithm \ref{main_alg} with AbsMax linkage, regardless of whether constraints are enforced or not, returns the same final clustering $\Pi^*_{\mathrm{AbsMax}}$ also returned by the Mutex Watershed Algorithm (MWS) \cite{wolf2018mutex}, which has empirical complexity $\mathcal{O}(N \log N)$.
\end{restatable}
% \begin{prop} 
% \end{prop}

\paragraph{Average, Single, and Complete Linkage} -- These three linkage criteria have been throughly studied on unsigned graphs, but never on signed graphs. In recent independent related work \cite{chehreghani2020hierarchical}, the authors prove that applying these three linkage methods to a signed graph is equivalent to applying them to the unsigned graph obtained by shifting all edge weights by a constant.  Here, we prove which of the algorithms studied here is ``intrinsically'' signed and do not have this invariance-property:
\begin{restatable}{prop}{invariantAlgs}
\label{prop:weight_shift_invariant}
We call an agglomerative algorithm ``weight-shift invariant'' if the dendrogram $T$ returned by the algorithm is invariant w.r.t. a shift of all edge weights $w_e$  by a constant $\alpha\in \mathbb{R}$. Among the variations of \algname{}, only hierarchical clustering with Average (HC-Avg), Single (HC-Single), and Complete linkage (HC-Complete) are weight-shift-invariant (see green box in Table~\ref{tab:linkage-criteria}).
\end{restatable}
\noindent To our knowledge, Average and Single linkage methods have never been combined with cannot-link constraints on signed graphs, so we name these algorithms \emph{HCC-Avg} and \emph{HCC-Single}. On the other hand, we prove that an agglomerative algorithm with Complete linkage (HC-Complete) returns the same clustering whether cannot-link-constraints are enforced or not (Lemma \ref{lemma:absMax_and_complete_property}, in Appendix).
% , we prove that with Complete linkage the agglomeration order does not change whether constraints are introduced or not.

@@

\paragraph{Algorithms defining an ultrametric} -- The connection between agglomerative algorithms and ultrametrics\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.} is well known. Usually, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. In our framework, a trivial ultrametric is always given by the height $h_T$ of the dendrogram. However, we now prove that, similarly to what has been done for HC on unsigned graphs \cite{johnson1967hierarchical,milligan1979ultrametric}, for some of the \algname{} variations we can define an ultrametric based on the edge weights and the interactions between clusters. To define this measure, we first map the signed interaction $\treeHeight{}$ defined in Eq.~\ref{eq:def_dendr_interact} to positive ``pseudo-distances'' $d_{T}:V \times V \rightarrow \mathbb{R}^{+}$:
\begin{align}\label{eq:UM_def}
d_{T}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv  \epsilon + \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v')
\end{align}
and where $\epsilon > 0$. 
% By definition, we have that $d_{T}(u,v)=d_{T}(v,u)$ and $d_{T}(u,u)= 0$. 
We then prove the following proposition:
\begin{restatable}{prop}{secondUltraMetricProperty}
\label{prop:ultraMetric2}
% Consider the binary rooted tree $T^*$ returned by the \TODO{all variants in Table 1} \algname{} Algorithm~\ref{main_alg}, given a choice of linkage criteria $\interact{}$ and the use (or not) of CannotLinkConstraints. 
Among the algorithms included in the \algname{} framework (see Table \ref{tab:linkage-criteria}), only Mutex Watershed and hierarchical clustering with Average (HC-Avg), Single (HC-Single) and Complete linkage (HC-Complete) define an ultrametric $(V, d_{T^*})$, where $d_{T^*}$ is defined in Eq.~\ref{eq:UM_def} and $T^*$ is the tree returned by the \algname{} Algorithm~\ref{main_alg}.
\end{restatable}
Now that we have defined the algorithms and characterized them from a theoretical point of view, we move to empirical evaluation in the next section.

### 2021-03-16 chapters/5_experiments_intro.tex
Shorten text

\emph{GASP: OurCNN + 3D-RAG + HCC-Avg} & \HollowBox &0.257 & 0.132 & 0.438& \textbf{0.063} \\  
\emph{GASP: OurCNN + 3D-RAG + HC-Avg} & \HollowBox &0.262 & 0.135 & 0.448 & \textbf{0.063}   \\  

@@

\textbf{Size of growing clusters: Sum vs Avg linkage} --
In all the studied clustering problems, we empirically observe that sum-linkage algorithms like GAEC tend to grow clusters one after the other, as shown in Fig.~\hyperref[fig:intro_figure]{\ref*{fig:intro_figure}d} and Fig.~\ref{fig:dendrograms} by the unique agglomeration order of GAEC\footnote{This \emph{flooding agglomeration-strategy} of GAEC was also observed in \cite{kardoostsolving}.}. This is intuitively explained by the  following: initially, many of the most attractive edge weights have very similar values; when the two nodes $u,v$ with the highest attraction are merged, there is a high chance that they will have a common neighboring node $t$ belonging to the same cluster; thus, the interaction between the merged nodes $uv$ and $t$ is likely assigned to the highest priority, because it is given by the sum of two highly attractive edge weights. This will then start a ``chain reaction'' where only a single cluster is agglomerated at the time. 
% This \emph{flooding strategy} in the agglomeration is not observed with other linkage methods, e.g. Average, which grow clusters of similar sizes in parallel.
In the following, we will show how this unique \emph{flooding strategy} of the sum-linkage methods can be both an advantage or a disadvantage, depending on the type of clustering problem.

@@

\textbf{Scores on CREMI instance-segmentation} -- 
SSBM graphs are \emph{non-planar}, and every edge has the same probability to be present in the graph. On the other hand, the \emph{gridGraph} and \emph{3D-RAG} graphs of Table~\ref{tab:datasets_and_energies} are sparse and have a very regular structure: regardless of whether a node represents a pixel or a superpixel, it will only have edge connections with its neighbors in the image (up to a certain hop distance). 
Tables~\ref{tab:scores_gridGraph}-\ref{tab:scores_3drag} show that average linkage methods (HC-Avg, HCC-Avg) strongly outperform other methods on \emph{CREMI-gridGraph} instances and also achieve the best scores on \emph{CREMI-3D-rag} graphs. Sum-based linkage methods (GAEC, HCC-Sum) have a two times higher ARAND error on grid-graphs and often return under-clustered segments (see failure cases in Fig.~\ref{fig:failure_cases}). This suggests that the \emph{flooding strategy} observed previously in the sum-linkage methods does not work on grid-graphs, because in this setup edge weights are predicted by a CNN and noise is strongly spatially-correlated \footnote{This effect is not as strong on \emph{3D-RAG} graphs, because edge weights are computed by averaging CNN predictions (and noise) over the boundaries of adjacent supervoxels.}.
To fully test this hypothesis, we conduct a set of experiments where the CNN predictions are perturbed by adding structured noise and simulating additional artifacts like ``holes'' in the boundary evidence\footnote{See Appendix \ref{sec:appendix_noise_gen} for details about how we perturbed the \emph{CREMI-gridGraph} problems by using Perlin noise \cite{perlin2001noise,perlin1985image}, which is one of the most common gradient noises used in procedural pattern generation.}. 
The plot in Fig.~\ref{fig:scores_structured_noise} confirms that HC-Avg and HCC-Avg perform well even in the presence of strong structured noise, followed by Sum-linkage algorithms and the Mutex Watershed algorithm (MWS). It is not a surprise that the AbsMax linkage used by MWS is not robust to this type of structured noise. However, the scores and runtimes in Table~\ref{tab:scores_gridGraph} prove how MWS can achieve high accuracy with considerably lower runtimes.

@@

\begin{figure}
\centering
% \includegraphics[width=0.48\textwidth,trim=300 100 300 100, clip]{./figs/dendrograms/new_agglo_order_OLD.png} % left bottom right top
\includegraphics[width=0.48\textwidth,trim=80 10 60 0, clip]{./figs/dendrograms/new_agglo_order.png} % left bottom right top
\caption{Clustering dynamics and accuracy of GASP variations on stochastic block models. The dendrograms result from three versions of \algname{} on a synthetic graph generated with SSBM ($250$ nodes, edge probability $p=0.05$, flipping probability $\eta=0.1$).  Red and blue colors show which of the two equal-sized ground-truth communities each node belongs to. At the top, dendrograms are truncated at the level of the final clustering $\Pi^*$ returned by \algname{}. \label{fig:dendrograms}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth,trim=90 10 60 0, clip]{./figs/comparison_GAEC_MWS_Avg.png} % left bottom right top
\caption{Failure cases of three versions of \algname{} applied to neuron segmentation. Only \emph{wrongly} segmented regions are highlighted in different warm colors. Red arrows point to wrongly split regions; yellow arrows point to false merge errors. HC-Avg returned the best segmentation. Data is 3D, hence the same color could be assigned to parts of segments that appear disconnected in 2D.  
\label{fig:failure_cases}}
\end{figure}

>>>>

\emph{GASP: OurCNN + 3D-RAG + HCC-Avg} & \CrossedBox &0.257 & 0.132 & 0.438& \textbf{0.063} \\  
\emph{GASP: OurCNN + 3D-RAG + HC-Avg} & \CrossedBox &0.262 & 0.135 & 0.448 & \textbf{0.063}   \\  

@@

\textbf{Size of growing clusters: Sum vs Avg linkage} --
In all the studied clustering problems, we empirically observe that sum-linkage algorithms like GAEC grow clusters one after the other, as shown in Fig.~\hyperref[fig:intro_figure]{\ref*{fig:intro_figure}d} and Fig.~\ref{fig:dendrograms} by the agglomeration order of GAEC\footnote{This \emph{flooding agglomeration-strategy} of GAEC was also observed in \cite{kardoostsolving}.}. This is intuitively explained by the  following: initially, many of the most attractive edge weights have very similar values; when the two nodes $u,v$ with the highest attraction are merged, there is a high chance that they will have a common neighboring node $t$ belonging to the same cluster; thus, the interaction between the merged nodes $uv$ and $t$ is likely assigned to the highest priority, because it is given by the sum of two highly attractive edge weights. This will then start a ``chain reaction'' where only a single cluster is agglomerated at the time. 
% This \emph{flooding strategy} in the agglomeration is not observed with other linkage methods, e.g. Average, which grow clusters of similar sizes in parallel.
In the following, we will show how this unique \emph{flooding strategy} of the sum-linkage methods can be both an advantage or a disadvantage, depending on the type of clustering problem.

@@

\textbf{Scores on CREMI instance-segmentation} -- 
SSBM graphs are \emph{non-planar}, and every edge has the same probability to be present in the graph. On the other hand, the \emph{gridGraph} and \emph{3D-RAG} graphs of Table~\ref{tab:datasets_and_energies} are sparse and have a very regular structure: regardless of whether a node represents a pixel or a superpixel, it will only have edge connections with its neighbors in the image (up to a certain hop distance). 
Tables~\ref{tab:scores_gridGraph}-\ref{tab:scores_3drag} show that average linkage methods (HC-Avg, HCC-Avg) strongly outperform other methods on \emph{CREMI-gridGraph} instances and also achieve the best scores on \emph{CREMI-3D-rag} graphs. Sum-based linkage methods (GAEC, HCC-Sum) have a two times higher ARAND error on grid-graphs and often return under-clustered segments (see failure cases in Fig.~\ref{fig:failure_cases}). This suggests that the \emph{flooding strategy} observed previously in the sum-linkage methods does not work on grid-graphs, because in this setup edge weights are predicted by a CNN and noise is strongly spatially-correlated \footnote{This effect is not as strong on \emph{3D-RAG} graphs, because edge weights are computed by averaging CNN predictions (and noise) over the boundaries of adjacent supervoxels.}.
To fully test this hypothesis, we conduct a set of experiments where the CNN predictions are perturbed by adding structured noise and simulating additional artifacts like ``holes'' in the boundary evidence\footnote{See Appendix \ref{sec:appendix_noise_gen} for details about how we perturbed the \emph{CREMI-gridGraph} problems by using Perlin noise \cite{perlin2001noise,perlin1985image}, which is one of the most common gradient noises used in procedural pattern generation.}. 
The plot in Fig.~\ref{fig:scores_structured_noise} confirms that HC-Avg and HCC-Avg are very robust algorihtms on this data, followed by Sum-linkage algorithms and the Mutex Watershed algorithm (MWS). It is not a surprise that the AbsMax linkage used by MWS is not robust to this type of structured noise. However, the scores and runtimes in Table~\ref{tab:scores_gridGraph} prove how MWS can achieve high accuracy with considerably lower runtimes.

@@

\begin{figure}
\centering
% \includegraphics[width=0.48\textwidth,trim=300 100 300 100, clip]{./figs/dendrograms/new_agglo_order_OLD.png} % left bottom right top
\includegraphics[width=0.48\textwidth,trim=0 10 60 0, clip]{./figs/dendrograms/new_agglo_order.png} % left bottom right top
\caption{Clustering dynamics and accuracy of GASP variations on stochastic block models. The dendrograms result from three versions of \algname{} on a synthetic graph generated with SSBM ($250$ nodes, edge probability $p=0.05$, flipping probability $\eta=0.1$).  Red and blue colors show which of the two equal-sized ground-truth communities each node belongs to. At the top, dendrograms are truncated at the level of the final clustering $\Pi^*$ returned by \algname{}. \label{fig:dendrograms}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth,trim=0 10 60 0, clip]{./figs/comparison_GAEC_MWS_Avg.png} % left bottom right top
\caption{Failure cases of three versions of \algname{} applied to neuron segmentation. Only \emph{wrongly} segmented regions are highlighted in different warm colors. Red arrows point to wrongly split regions; yellow arrows point to false merge errors. HC-Avg returned the best segmentation. Data is 3D, hence the same color could be assigned to parts of segments that appear disconnected in 2D.  
\label{fig:failure_cases}}
\end{figure}

### 2021-03-16 chapters/6_conclusions.tex
Shorten text

% !TEX root = ../agglo_clust_review.tex
\section{Conclusion}
We have presented a unifying framework for agglomerative clustering of graphs with both positive and negative edge weights; and we have shown that some existing clustering algorithms, e.g. the Mutex Watershed, can be reformulated as special cases of this framework. 
We have analyzed several theoretical and empirical properties of the algorithms included in this framework. On instance segmentation, algorithms based on an average linkage criterion outperformed all the others: they proved to be simple and remarkably robust approaches to process short- and long-range predictions of a CNN.
 % applied to an instance segmentation task.
On biological images, these simple average agglomeration algorithms achieve state-of-the-art results without requiring the user to spend much time tuning complex task-dependent pipelines based on super-pixels.
% can represent a valuable choice for a user who is not willing to spend much time tuning complex task-dependent pipelines based on super-pixels.  
% In future work, we plan to explore common theoretical properties of the algorithms included in the framework.
% On biological images, this simple average agglomeration algorithm superpixels

>>>>

% !TEX root = ../agglo_clust_review.tex
\section{Conclusion}
We have presented a unifying framework for agglomerative clustering of graphs with both positive and negative edge weights; and we have shown that some existing clustering algorithms, e.g. the Mutex Watershed, can be reformulated as special cases of this framework. 
We have analyzed several theoretical and empirical properties of the algorithms included in this framework. On instance segmentation, algorithms based on an average linkage criterion outperformed all the others: they proved to be simple and robust approaches to process short- and long-range predictions of a CNN.
 % applied to an instance segmentation task.
On biological images, these simple average agglomeration algorithms achieve state-of-the-art results without requiring the user to spend much time tuning complex task-dependent pipelines based on super-pixels.
% can represent a valuable choice for a user who is not willing to spend much time tuning complex task-dependent pipelines based on super-pixels.  
% In future work, we plan to explore common theoretical properties of the algorithms included in the framework.
% On biological images, this simple average agglomeration algorithm superpixels
